Comparison of Terrain Following and Cut Cell Grids using a Non-Hydrostatic Model

### 2016-01-09 src/tf-cutcell-comparison/responses2.tex
add responses for resting atmosphere test

\begin{quotation}
 Lines 131-147 on cut-cell grid generation - It's still not clear how or why a cell vertex  is moved laterally in constructing the cut-cell mesh. Taking Fig. 2 as an example, it  seems that following the authors' directions to move all vertices lying underneath the  orography on the uniform grid up to the surface would produce four grid cells as  shown in the figure below rather than the three cells shown in Fig. 2. This four-cell structure appears to be consistent with the generation of thin cells as the authors  point out in Fig. 5c. If lateral displacement of a vertex is allowed, how is it decided  whether it should be moved and how is it decided where along the surface it should  be located (i.e. the location of point v in Fig. 2)?
\end{quotation}
We have found an alternative method of generating meshes that does not require the use of the \texttt{snappyHexMesh} utility.  This new method does not move vertices laterally and is more straightforward to describe.  This new method creates meshes that retain the long, thin cells from the old method.  The new method is described in section 2 and figure 2.  All test results on cut cell grids have been updated.  \TODO{note changes that have occurred: centred linear scheme no longer unstable in TF advection; larger errors in thermal advection test?}

\begin{quotation}
	Section 3b - Finite difference linear advection model - This advection model is solving the same flux form equation (4) as in the previous subsection. Thus, the advection operator in (6) should be in terms of $\partial (u\phi)/\partial x$.  The finite differencing in (6) can also be considered a finite-volume representation, rewriting it as 
	\begin{align*}
		\frac{\partial u \phi}{\partial x} &= \frac{1}{\Delta x} \left( u_{i+1/2} F_{i+1/2} - u_{i - 1/2} F_{i - 1/2}\right), \\
		F_{i+1/2} &= \frac{1}{12} \left( - \phi_{i+2} + 7 \phi_{i+1} + 7 \phi_i - \phi_{i-1} \right).
	\end{align*}
\end{quotation}
Amended as suggested, many thanks.

\begin{quotation}
 Lines 265-267 and table 1 – The authors are unnecessarily complicating their tests  by comparing results from the cut-cell grid and a regular grid. These grids should be identical for this test and merit no further comparative consideration. There is no need to simulate with both grids, show their identical behavior in table 1, and then  state that the same behavior ``is to be expected.''
\end{quotation}
\TODO{agreed, although the comparison with the fourth order scheme in table 1 is perhaps useful?  I have not removed it yet.}

\begin{quotation}
 Lines 271-279 and Table 1 – The $\ell_2$ error norms have changed by more than an order  of magnitude from those reported in the original manuscript. What changed?
\end{quotation}
In the first draft, the $\ell_2$ values had not been normalised, such that $\ell_2 = \sqrt{\sum_c \left(\phi - \phi_T\right)^2 \mathcal{V}_c}$.  This was fixed in the second draft so that $\ell_2$ was calculated correctly according to equation (13).
\begin{quotation}
 Lines 280-285 – Reviewer 2 questioned the large change in the $\ell_2$ error norm when 
 the domain width was changed from 301 to 300 km. I share the concern expressed 
 by this reviewer and find the authors' added explanation (``It's likely that changes in 
 the domain width affect the wave power spectrum of the discrete terrain profile.'') to 
 be inadequate. If the results are sensitive to the exact location of the lateral 
 boundaries, then the numerical experiment is not well designed; the boundaries 
 should be moved far enough away from the mountain so that this sensitivity 
 disappears.  Is it possible that the authors shifted the location of the terrain by a half 
 grid interval when they changed the domain width by one grid interval? This would 
 occur if they just defined the center of the terrain profile to be in the center of the 
 domain. If that’s the case, then the results are understandable since numerically, it's a 
 different terrain shape. However, the changing behavior would have nothing to do 
 with the domain width, and thus this sensitivity test should then be removed from 
 the paper altogether. One would get the same sensitivity leaving the domain width 
 fixed and just shifting the terrain location by one-half grid interval.  
\end{quotation}
Yes, the change in domain width causes the terrain to be shifted by $\Delta x / 2$, and it is this which results in a change in accuracy.  We had not intended to perform a sensitivity test, but only to clarify the specification for the \citet{schaer2002} advection test.

@@

\begin{quotation}
 Lines 321-322 – When the flow is aligned with the grid there is no cross coordinate 
 flow and therefore the numerics are essentially reduced to a one-dimensional 
 advection equation. I don't see how grid distortion is even an issue. 
\end{quotation}
The advection scheme fits a multidimensional polynomial using a stencil of cell centroids, hence grid distortions could affect the resulting least squares fit.  If the test used a dimensionally-split advection scheme with a terrain-following coordinate transform, then the test would indeed reduce to a one-dimensional problem.  We have provided a more detailed motivation of this test in the introductory paragraph starting at line \TODO{xxx}.

\begin{quotation}
 Lines 364-368 – The authors did not need to run these resting atmosphere 
 simulations to reach the conclusion stated in this paragraph. As stated in my first 
 review, this test case does not challenge the numerics on a cut-cell grid since the horizontal pressure-gradient errors will obviously be reduced in the vicinity of the 
 low level inversion where the coordinates surfaces are perfectly horizontal. A better 
 test case might be to specify an inversion layer that intersects the terrain, since on 
 the cut-cell grid, it's only just adjacent to the terrain where pressure gradient 
 calculations between cell centers are not horizontal.   
\end{quotation}
\TODO{good idea.  Should we abandon the klemp2011 test entirely or just put less emphasis on its results?}

\begin{quotation}
	Section 4d – Sch\"{a}r mountain-wave tests. These tests are more interesting now that 
 the case has been simulated with a variety of grid resolutions. However, I'm 
 surprised that the analyses of these results seem to focus on behavior far 
 downstream of the mountain, as is evident in Figs. 6 and 7. I'm not aware of any 
 published results that extend beyond x = 25 km in Fig. 6. Since the amplitude of the 
 disturbances are decaying with downstream distance, small inaccuracies in behavior 
 of the absorbing layer and boundary conditions have larger impact such that the 
 relative errors in comparison to the analytic solution become large. (Note, for 
 example the erroneous extension of negative perturbation theta contours to the 
 surface at x = 25 km in Fig. 6.)  All we learn from Fig. 7b is that errors in theta near 
 the surface in the cut-cell cases, which must have been produced in the vicinity of the 
 terrain, continue to be swept downstream. I think it would be more informative to 
 focus on solution differences between the BTF and cut-cell results near the terrain as 
 is done in Fig. 8 for the terrain-following theta advection. The authors could display 
 these difference fields for the various grid resolutions in exactly the same manner as 
 in Fig. 8. The theta advection test presented by the authors is a very nice complement 
 to the gravity-wave tests. It confirms that the BTF solution is behaving properly and 
 therefore lends support to interpreting differences in the two grids as inaccuracies in 
 the cut cell treatment. I would expect that the difference fields in the Sch\"{a}r case are 
 at least qualitatively similar to those exhibited in the theta advection case, which 
 would further confirm the relevance of this advection test as well as the conclusion 
 that errors in the advection of theta through the cut cells may be responsible for the 
 anomalies appearing with this grid.
\end{quotation}
We considered plotting heatmaps but found that it was not easy to distinguish between gravity waves and numerical errors.  The position of the sample line was chosen so that numerical errors would be revealed more clearly: far away from the mountain, the wave amplitude is small but the potential temperature errors remain constant in the lee of the mountain.  The manuscript now explains this choice at line \TODO{xxx}.

\begin{quotation}
 Lines 479-480 – As mentioned above, for the horizontal advection test, shouldn't the  results for the cut-cell and regular grids be identical?
\end{quotation}
This was a mistake: table 1 confirmed that min, max and $\ell_2$ errors were identical on both grids.  This line now reads, `in the horizontal advection test, accuracy was greatest on the cut cell grid'.

\begin{quotation}
 Lines 481-482 – Emphasizing an error reduction of two orders of magnitude for this 
 resting atmosphere case with cut cells compared to a terrain-following coordinate is 
 not very informative. In both cases the errors appear to be quite small, and thus the 
 more relevant question is whether these differences are of any realistic significance. 
\end{quotation}
\TODO{this conclusion may change when we run the new resting atmosphere test}

>>>>

\begin{quotation}
 Lines 131-147 on cut-cell grid generation -- It's still not clear how or why a cell vertex  is moved laterally in constructing the cut-cell mesh. Taking Fig. 2 as an example, it  seems that following the authors' directions to move all vertices lying underneath the  orography on the uniform grid up to the surface would produce four grid cells as  shown in the figure below rather than the three cells shown in Fig. 2. This four-cell structure appears to be consistent with the generation of thin cells as the authors  point out in Fig. 5c. If lateral displacement of a vertex is allowed, how is it decided  whether it should be moved and how is it decided where along the surface it should  be located (i.e. the location of point v in Fig. 2)?
\end{quotation}
We have found an alternative method of generating meshes that does not require the use of the \texttt{snappyHexMesh} utility.  This new method does not move vertices laterally and is more straightforward to describe.  This new method creates meshes that retain the long, thin cells from the old method.  The new method is described in section 2 and figure 2.  All test results on cut cell grids have been updated.

\begin{quotation}
	Section 3b -- Finite difference linear advection model - This advection model is solving the same flux form equation (4) as in the previous subsection. Thus, the advection operator in (6) should be in terms of $\partial (u\phi)/\partial x$.  The finite differencing in (6) can also be considered a finite-volume representation, rewriting it as 
	\begin{align*}
		\frac{\partial u \phi}{\partial x} &= \frac{1}{\Delta x} \left( u_{i+1/2} F_{i+1/2} - u_{i - 1/2} F_{i - 1/2}\right), \\
		F_{i+1/2} &= \frac{1}{12} \left( - \phi_{i+2} + 7 \phi_{i+1} + 7 \phi_i - \phi_{i-1} \right).
	\end{align*}
\end{quotation}
Amended as suggested, many thanks.

\begin{quotation}
 Lines 265-267 and table 1 -- The authors are unnecessarily complicating their tests  by comparing results from the cut-cell grid and a regular grid. These grids should be identical for this test and merit no further comparative consideration. There is no need to simulate with both grids, show their identical behavior in table 1, and then  state that the same behavior ``is to be expected.''
\end{quotation}
We have removed the discussion but we believe that the inclusion of results on the regular grid is still useful because it enables a comparison with the fourth order results from \citet{schaer2002} which are not available on the cut cell grid.

\begin{quotation}
 Lines 271-279 and Table 1 -- The $\ell_2$ error norms have changed by more than an order  of magnitude from those reported in the original manuscript. What changed?
\end{quotation}
In the first draft, the $\ell_2$ values had not been normalised, such that $\ell_2 = \sqrt{\sum_c \left(\phi - \phi_T\right)^2 \mathcal{V}_c}$.  This was fixed in the second draft so that $\ell_2$ was calculated correctly according to equation (13).
\begin{quotation}
 Lines 280-285 -- Reviewer 2 questioned the large change in the $\ell_2$ error norm when 
 the domain width was changed from 301 to 300 km. I share the concern expressed 
 by this reviewer and find the authors' added explanation (``It's likely that changes in 
 the domain width affect the wave power spectrum of the discrete terrain profile.'') to 
 be inadequate. If the results are sensitive to the exact location of the lateral 
 boundaries, then the numerical experiment is not well designed; the boundaries 
 should be moved far enough away from the mountain so that this sensitivity 
 disappears.  Is it possible that the authors shifted the location of the terrain by a half 
 grid interval when they changed the domain width by one grid interval? This would 
 occur if they just defined the center of the terrain profile to be in the center of the 
 domain. If that’s the case, then the results are understandable since numerically, it's a 
 different terrain shape. However, the changing behavior would have nothing to do 
 with the domain width, and thus this sensitivity test should then be removed from 
 the paper altogether. One would get the same sensitivity leaving the domain width 
 fixed and just shifting the terrain location by one-half grid interval.  
\end{quotation}
Yes, the change in domain width causes the terrain to be shifted by $\Delta x / 2$, and it is this which results in a change in accuracy.  We had not intended to perform a sensitivity test, but only to clarify the specification for the \citet{schaer2002} advection test.

@@

\begin{quotation}
 Lines 321-322 -- When the flow is aligned with the grid there is no cross coordinate 
 flow and therefore the numerics are essentially reduced to a one-dimensional 
 advection equation. I don't see how grid distortion is even an issue. 
\end{quotation}
The advection scheme fits a multidimensional polynomial using a stencil of cell centroids, hence grid distortions could affect the resulting least squares fit.  If the test used a dimensionally-split advection scheme with a terrain-following coordinate transform, then the test would indeed reduce to a one-dimensional problem.  We have provided a more detailed motivation of this test in the introductory paragraph starting at line \TODO{xxx}.

\begin{quotation}
 Lines 364-368 -- The authors did not need to run these resting atmosphere 
 simulations to reach the conclusion stated in this paragraph. As stated in my first 
 review, this test case does not challenge the numerics on a cut-cell grid since the horizontal pressure-gradient errors will obviously be reduced in the vicinity of the 
 low level inversion where the coordinates surfaces are perfectly horizontal. A better 
 test case might be to specify an inversion layer that intersects the terrain, since on 
 the cut-cell grid, it's only just adjacent to the terrain where pressure gradient 
 calculations between cell centers are not horizontal.   
\end{quotation}
We have now included test results with an inversion layer that intersects the terrain, which have similar maximum vertical velocity magnitudes on terrain following and cut cell grids.  We have also included results for the original test to enable comparison with \citet{klemp2011}.

\begin{quotation}
	Section 4d -- Sch\"{a}r mountain-wave tests. These tests are more interesting now that 
 the case has been simulated with a variety of grid resolutions. However, I'm 
 surprised that the analyses of these results seem to focus on behavior far 
 downstream of the mountain, as is evident in Figs. 6 and 7. I'm not aware of any 
 published results that extend beyond x = 25 km in Fig. 6. Since the amplitude of the 
 disturbances are decaying with downstream distance, small inaccuracies in behavior 
 of the absorbing layer and boundary conditions have larger impact such that the 
 relative errors in comparison to the analytic solution become large. (Note, for 
 example the erroneous extension of negative perturbation theta contours to the 
 surface at x = 25 km in Fig. 6.)  All we learn from Fig. 7b is that errors in theta near 
 the surface in the cut-cell cases, which must have been produced in the vicinity of the 
 terrain, continue to be swept downstream. I think it would be more informative to 
 focus on solution differences between the BTF and cut-cell results near the terrain as 
 is done in Fig. 8 for the terrain-following theta advection. The authors could display 
 these difference fields for the various grid resolutions in exactly the same manner as 
 in Fig. 8. The theta advection test presented by the authors is a very nice complement 
 to the gravity-wave tests. It confirms that the BTF solution is behaving properly and 
 therefore lends support to interpreting differences in the two grids as inaccuracies in 
 the cut cell treatment. I would expect that the difference fields in the Sch\"{a}r case are 
 at least qualitatively similar to those exhibited in the theta advection case, which 
 would further confirm the relevance of this advection test as well as the conclusion 
 that errors in the advection of theta through the cut cells may be responsible for the 
 anomalies appearing with this grid.
\end{quotation}
We considered plotting heatmaps but found that it was not easy to distinguish between gravity waves and numerical errors.  The position of the sample line was chosen so that numerical errors would be revealed more clearly: far away from the mountain, the wave amplitude is small but the potential temperature errors remain constant in the lee of the mountain.  The manuscript now explains this choice at line \TODO{xxx}.

\begin{quotation}
 Lines 479-480 -- As mentioned above, for the horizontal advection test, shouldn't the  results for the cut-cell and regular grids be identical?
\end{quotation}
This was a mistake: table 1 confirmed that min, max and $\ell_2$ errors were identical on both grids.  This line now reads, `in the horizontal advection test, accuracy was greatest on the cut cell grid'.

\begin{quotation}
 Lines 481-482 -- Emphasizing an error reduction of two orders of magnitude for this 
 resting atmosphere case with cut cells compared to a terrain-following coordinate is 
 not very informative. In both cases the errors appear to be quite small, and thus the 
 more relevant question is whether these differences are of any realistic significance. 
\end{quotation}
The advantage of the cut cell grid is not seen when the inversion layer is lowered to intersect the terrain.  We have removed the emphasis on the error reduction from the concluding remarks of section 4c.
